---
output: github_document
---

NOTE: This package is under development, and a few functions are undergoing a re-write. Additionally, more complete documentation and test cases are in progress.

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# icicle

<!-- badges: start -->
<!-- badges: end -->

The goal of icicle is to calculate and tidy ice cycle variables from flow data collected by the Water Survey of Canada. This package provides a set of functions that calculate the freeze/thaw dates, onset of freshet, and length of continuous ice coverage for each year. Additional tidying functions remove years from the dataframe with consecutive NA values, and add a column of "day of the year", or Julian dates. 

## Installation

You can install the development version of icicle from [GitHub](https://github.com/) with:

``` r
#install.packages("devtools")
devtools::install_github("Jacqui-123/icicle")
```

## Example

This is a basic example of how you might use the functions in this package:

```{r example}
library(icicle)
## tbd example 
```


```{r, include=FALSE, echo = FALSE, warning = FALSE}

#You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.


#In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

```

