---
output: github_document
---

NOTE: This package is under development, and a few functions are undergoing a re-write. More complete documentation and test cases are in progress.

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%")
```

# icicle

<!-- badges: start -->
<!-- badges: end -->

The goal of icicle is to calculate and tidy ice cycle variables from flow data collected by the Water Survey of Canada. This package provides a set of functions that calculate the freeze/thaw dates, onset of freshet, and length of continuous ice coverage for each year. Additional tidying functions remove years from the dataframe with consecutive NA values, and add a column for "day of the year". 

## Installation

You can install the development version of icicle from [GitHub](https://github.com/) with:

``` r
#install.packages("devtools")
devtools::install_github("Jacqui-123/icicle")
```

## Example

This is a basic example of how you might use the functions in this package:

```{r, ex1}
#import data for a water station from the Water Survey of Canada using tidyhydat::hy_daily_flows()
#here we use a test data file:
data_example <- read.csv("testdata.csv")

```

```{r, echo= FALSE}
library(knitr)
kable(data_example[1:4,], caption = " ") 
```

Add a Julian date, or "day of the year" column to the data frame 
```{r ex2}
data_doy <- icicle::ice_day_of_wyear(data_example)
```

```{r, echo= FALSE}
library(knitr)
kable(data_doy[1:4,], caption = " ") 

```

Calculate the length of continuous ice coverage using the "B" dates marked on the Water Survey of Canada's database 

```{r, echo = FALSE, message=FALSE, warning=FALSE}
#tidy and format data
library(tidyverse)
data_doy <- data_doy %>%
  dplyr::mutate(waterYear = as.character(waterYear)) %>%
  dplyr::mutate(Date = as.Date(Date)) 

```

```{r ex3, warning=FALSE, message = FALSE, fig.align='center'}
library(icicle)
results_g1 <- ice_group_1(data_doy)
```


```{r, echo = FALSE}
library(knitr)
results_1 <- results_g1 %>% filter("Year" != "2021")
kable(results_g1[1:4,], title = "Ice Coverage Results") 

```
Plot the length of continuous ice coverage per year to visualize change over time

```{r, echo= FALSE, warning=FALSE}
#Plot the results of ice coverage for each year
#plot(x= results_g1$Year, y = results_g1$Ice_coverage, pch = 16,
 #  cex = 1.0, xlab = "Year", ylab = "Days of Continuous Ice Cover")
library(ggplot2)
library(tidyverse)
library(caTools)
data <- results_g1 %>% filter(Year != "2021")
ggplot(data, aes(x=Year, y=Ice_coverage)) +
    geom_point(col= "blue") +     # Use hollow circles
  scale_x_discrete(breaks = seq(1960, 2020, 10)) +
  ylab("Days of continuous ice coverage") +
  ylim(40, 200)+
  theme_bw() 


```

Calculate the freeze and thaw dates for each year using the onset of continuous "B" dates 

```{r ex5}
library(icicle)
results_g2 <- ice_group_2(data_doy)
```


```{r , echo = FALSE}
library(knitr)
kable(results_g2[7:9,], title = "Results Table for Freeze and Thaw Dates") 

```
```

```{r, include=FALSE, echo = FALSE, warning = FALSE}

#You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

#In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

```

