---
output: github_document
---

NOTE: This package is under development, and a few functions are undergoing a re-write. Additionally, more complete documentation and test cases are in progress.

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# icicle

<!-- badges: start -->
<!-- badges: end -->

The goal of icicle is to calculate and tidy ice cycle variables from flow data collected by the Water Survey of Canada. This package provides a set of functions that calculate the freeze/thaw dates, onset of freshet, and length of continuous ice coverage for each year. Additional tidying functions remove years from the dataframe with consecutive NA values, and add a column for "day of the year". 

## Installation

You can install the development version of icicle from [GitHub](https://github.com/) with:

``` r
#install.packages("devtools")
devtools::install_github("Jacqui-123/icicle")
```

## Example

This is a basic example of how you might use the functions in this package:

```{r ex1}

library(icicle)
#import data from tidyhydat using tidyhydat::hy_daily_flows()
#here we use a prepared test data file
data_example <- read.csv("testdata.csv")

```

Add a Julian date, or "day of the year" to the data frame 
```{r ex2}
library(icicle)
#add a day of the year
data_doy <- ice_day_of_wyear(data_example)
```


```{r, echo= FALSE}
library(knitr)
kable(data_doy[1:6,], caption = " ") 

```

Calculate the length of continuous ice coverage using the "B" dates marked on the Water Survey of Canada's database 
```{r ex3, warning=FALSE, message = FALSE}
library(icicle)
library(tidyverse)

#tidy and format data
data_doy <- data_doy %>%
  dplyr::mutate(waterYear = as.character(waterYear)) %>%
  dplyr::mutate(Date = as.Date(Date)) 

results_g1 <- ice_group_1(data_doy)
```


```{r, echo = FALSE}
library(knitr)
kable(results_g1[1:3,], caption = " ") 

```

Plot the results of ice coverage for each year
```{r, echo= FALSE}

plot(results_g1)

```

Calculate the freeze and thaw dates for each year using the onset of continuous "B" dates 

```{r ex5}
library(icicle)
results_g2 <- ice_group_2(data_doy)
```


```{r , echo = FALSE}
library(knitr)
kable(results_g2[7:9,], caption = " ") 

```
```

```{r, include=FALSE, echo = FALSE, warning = FALSE}

#You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

#In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

```

